---
title: "Magrittr Workshop"
date: "2018-11-14"
output:
   html_document:
     theme: yeti
     toc: true
     toc_float: true
     number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
```



```{r}
library(magrittr)
library(ggplot2) # or library(tidyverse)
```


Start by having a look at the CO2 dataset:

```{r}
head(CO2)
```

What we would like to do now, is to make a scatter plot of he data.
But we only want to plot plants from Quebec, with "nonchilled" treatment.

Basically we want to:
* Take our CO2 dataset -

* Subset the Quebec plant

* Subset the nonchilled treatment

* Plot the data


Written in oneline this would be written:

```{r}
ggplot(subset(subset(CO2, Type=="Quebec"), Treatment=="nonchilled"), aes(x=conc, y=uptake)) +
         geom_point()

```


The code can be improved by:

```{r}

Quebec_CO2 <- subset(CO2, Type=="Quebec")
nonchilled_Quebec_CO2 <- subset(Quebec_CO2, Treatment=="nonchilled")
ggplot(nonchilled_Quebec_CO2, aes(x=conc, y=uptake)) +
  geom_point()

```


But in this case, we need to create lots of intermediates.
Try to use pipes %>%  to create the same plot with one single piece of code.

Remember what we want to achieve:

* Take our CO2 dataset -

* Subset the Quebec plant

* Subset the nonchilled treatment

* Plot the data



```{r, echo=FALSE, eval=FALSE}
library(magrittr)
library(ggplot2)
CO2 %>%
  subset(Type=="Quebec") %>% 
  subset(Treatment=="nonchilled") %>% 
  ggplot(aes(x=conc, y=uptake)) +
  geom_point()
```

