---
title: "R Functions Workshop 1"
author: Andreas Novotny
date: "2018-11-14"
output:
   html_document:
     theme: yeti
     toc: true
     toc_float: true
     number_sections: true
     code_folding: "hide"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T,
                      results = "hide")
```

This exercise will guide you through how to write your own R functions and how to use it.

After each assignment there is a hidden code chunk, containing clues for how to solve the problem. Start by trying on your own. If you are stuck, click on the "code" button.


# "Hello R"

We will now start by writing our very first function, and then executing it.

[CONTINUE TO ADVANCED EXCERCISE 2](functionsexc2.html)

1. Create a function (without any input) called "helloR" that prints "Hello World" when executed.

**Create function**

*Hint:*
```{r}
# Remember the basic syntax for designing a function?
# If not, repeat this:

name_of_function <- function() {
  
}
```

*Solution:*

```{r}
helloR <- function() {
  
  print("Hello World")

}
```


**Execute function**

*solution:*

```{r}
helloR()
```


2. Well done, not that tricky right? Now lets ad some option to the function. If we provide a name as input, perhaps it can print "Hello NNN" ? Lets try!


*Hint:*

```{r}
# Remember where we shoud put the input?

name_of_function <- function(input) {
  
}
```

*Hint:*

```{r}
# Ahaaa, how to combine Hello and XXX?
# Perhaps this will give you a clue:

name <- "Andreas"
print(paste("Hello", name))
```

*Solution:*

```{r}
#Define function

helloR <- function(name) {
  
  print(paste("Hello", name))

}

# Execute function

helloR("Andreas")
helloR("Malie")
```

3. Cool!!! We made it!!! Now we have a full function with input, and output. But, as we are going,
lets define a default input, if no other input is given. Lets put your own name if no other input is given.

*Hint:*
```{r}
# Remember how to define default inputs, no?

name_of_function <- function(input=Default) {
  
}
```

*Solution:*

```{r}
#Define function

helloR <- function(name="Andreas") {
  
  print(paste("Hello", name))
}

# Execute function - try some different options:

helloR()
helloR("Andreas")
helloR("Malie")
```


**Nice! We are done!**

Now lets see if any of your mates need help - and the presentation will continue shortly!





# Normalization function

Usually, we dont want R to just print things to the screen. We usually want to store the output as a new variable, that we can later use for other calculations.

1. Now, create a function that takes as input a vector of values (i.e a column from a data.frame) and returns a a vector where all the values has been normalized towards the mean (xi = xi/mean(x)). We can use the "Iris" dataset as example. The function can be called "normalizeR"

*Hint 1:*

```{r}
# Lets repeat the syntax:

function_name <- function(input=default) {
  
  output <- "make something out of input"
  
  return(output)
}
```

*Hint 2:*

```{r}
# What should we do with the input?
normalized <- iris$Sepal.Length / mean(iris$Sepal.Length)
```

*Solution:*

```{r, eval= FALSE}
#Now, just put it together:

normalizeR <- function(inputvector) {
  
  normalized <- inputvector/mean(inputvector)
  
  return(normalized)
}

# Lets try the function out:

normalizeR(iris$Sepal.Length)
# Thets compare the output
boxplot(iris$Sepal.Length)
boxplot(normalizeR(iris$Sepal.Length))
```


2. Now, we would like to communicate a bit with the user. What happens if the user do not provide a vector? Make the function stop and provide an error message.

*Hint 1:*

```{r, eval=FALSE}
# The stop function
stop("Write an error message here")
```

*Hint 2:*
```{r}
# Do you remember the ifelse statements?

if (class(iris$Sepal.Length) != "numeric") {
  "Take some action"
}

```


*Solution:*

```{r, eval= FALSE}
#Pay attention to the use of the braces and indentations:

normalizeR <- function(inputvector) {
  
  if (class(inputvector) != "numeric") {
    stop("Input is not a numeric vector")
    }
  
  normalized <- inputvector/mean(inputvector)
  
  return(normalized)
}


# Try your function:

normalizeR(iris$Sepal.Length)
normalizeR("Johansson")
```


# Wrapper functions

Have you not always been annoyed by the default black color in Rs plot functions? Soo boring! Lets hot it up. Create a function that per default plots anything in red colors. Call the function redplot


*Hint:*

```{r, eval=FALSE}
#This is how a plot would look like:

plot(x=iris$Sepal.Length, y=iris$Sepal.Width, col="red")
```

*Solution:*

```{r}
redplot <- function(x, y, col="red") {
  plot(x=x, y=y, col=col)
}
```


But hey! What if the user wants to add i.e. a title on the x-axis? Leave roome in the formula for other arguments in the plot function:

*Solution:*

```{r, eval=FALSE}
redplot <- function(x, y, col="red", ...) {
  plot(x=x, y=y, col=col, ...)
}


# Ore make it eaven nicer:

redplot <- function(..., col="red") {
  plot(..., col=col)
}

# Try it out

redplot(x=iris$Sepal.Length, y=iris$Sepal.Width)
redplot(x=iris$Sepal.Length, y=iris$Sepal.Width, xlab="Sepal Length", ylab="Sepal Width")
```

Its quite marvelous, is it not? See if you can use your function for other purposes.


[CONTINUE TO ADVANCED EXCERCISE 2](functionsexc2.html)


# Session Info

These packages were used during this session:

```{r, results="show"}
sessionInfo()
```






