---
title: "R Functions Workshop 2"
author: Andreas Novotny
date: "2018-11-14"
output:
   html_document:
     theme: yeti
     toc: true
     toc_float: true
---

[BACK TO EXCERCISE 1](functionsexc.html)


We will now have a look at a dataset created during a basic ecology course for bachelor students.
The data describe feeding- and survival rates of Daphnia from two different lakes, during two different temperature conditions. The full code of the students data analysis is pasted below.

[Download the dataset here](https://github.com/deepskillsr/Data_for_Students/blob/master/dafnia.rds)




```{r, results="show"}
dafnia <- readRDS("dafnia.rds")
head(dafnia)
```



### Feeding Rate

1. Create data summary
```{r, results="show"}
SumTable <- aggregate(dafnia$Feeding,
                      by = list(Lake = dafnia$Lake,
                                Temp = dafnia$Temp),
                      FUN = function(x) c(mean = mean(x),
                                          sd = sd(x),
                                          n = length(x)))
head(SumTable)
```

2. Format the table
```{r, results="show"}
SumTable <- do.call(data.frame, SumTable)
SumTable$se <- SumTable$x.sd / sqrt(SumTable$x.n)
colnames(SumTable) <- c("Lake", "Temp", "mean", "sd", "n", "se")
SumTable
```

3. Plot data
```{r, results="show"}
library("ggplot2")

ggplot(data = SumTable, aes(x = factor(Temp),
                            y = mean, fill = factor(Lake)))+
  geom_bar(stat = "identity",
           position = position_dodge(width =0.9)) +
  geom_errorbar(aes(ymax = SumTable$mean+SumTable$se,
                    ymin = SumTable$mean-SumTable$se),
                position = position_dodge(width = 0.9),
                width = 0.25) +
  labs(x = "Temperatur (grC)", y = "Ätna Celler") +
  scale_fill_manual(values=c("royalblue1", "peachpuff1"), name="Sjö")
```

### Survival Rate

1. Create data summary
```{r, results="show"}
SumTable <- aggregate(dafnia$Survival,
                      by = list(Lake = dafnia$Lake,
                                Temp = dafnia$Temp),
                      FUN = function(x) c(mean = mean(x),
                                          sd = sd(x),
                                          n = length(x)))
head(SumTable)
```

2. Format the table
```{r, results="show"}
SumTable <- do.call(data.frame, SumTable)
SumTable$se <- SumTable$x.sd / sqrt(SumTable$x.n)
colnames(SumTable) <- c("Lake", "Temp", "mean", "sd", "n", "se")
SumTable
```

3. Plot data
```{r, results="show"}
library("ggplot2")

ggplot(data = SumTable, aes(x = factor(Temp),
                            y = mean, fill = factor(Lake)))+
  geom_bar(stat = "identity",
           position = position_dodge(width =0.9)) +
  geom_errorbar(aes(ymax = SumTable$mean+SumTable$se,
                    ymin = SumTable$mean-SumTable$se),
                position = position_dodge(width = 0.9),
                width = 0.25) +
  labs(x = "Temperatur (grC)", y = "Överlevande") +
  scale_fill_manual(values=c("royalblue1", "peachpuff1"), name="Sjö")
```


# 1. Make code nicer by structuring the code in functions.

As you can see, the pretty long piece of code is repeated two times. Is it possible to restructure the code in functions, to avoid repetition, and make the code smaller?



# 2. Custumize the function

As a last step, try to generalize this function as much as you can. Can you make a function,
or a set of function that creates any kind of bar plot in ggplot? Lets see how far we can go!


[BACK TO EXCERCISE 1](functionsexc.html)





# 3. Possible solution:


```{r}
sumtable <- function(data = dafnia, y = "Feeding", x1 = "Lake", x2 = "Temp") {
  SumTable <- aggregate(data[[y]],
                      by = list(x1 = data[[x1]],
                                x2 = data[[x2]]),
                      FUN = function(x) c(mean = mean(x),
                                          sd = sd(x),
                                          n = length(x)))
  
  SumTable <- do.call(data.frame, SumTable)
  SumTable$se <- SumTable$x.sd / sqrt(SumTable$x.n)
  colnames(SumTable) <- c(x1, x2, "mean", "sd", "n", "se")
  return(SumTable)
  
}
```


```{r}
plotbar <- function(data = dafnia, y = "Feeding", x = "Lake", fill = "Temp", ...) {
  require("ggplot2")
  data <-sumtable(data = data, y = y, x1 = x, x2 = fill)
  plot <- ggplot(data, aes(x = factor(data[[x]]),
                              y = mean, fill = factor(data[[fill]])))+
    geom_bar(stat = "identity",
             position = position_dodge(width =0.9), ...) +
    geom_errorbar(aes(ymax = data$mean+data$se,
                      ymin = data$mean-data$se),
                  position = position_dodge(width = 0.9),
                  width = 0.25) +
    labs(x = x, y = y) +
    scale_fill_manual(values = c("royalblue1", "peachpuff1"), name = fill)
    
  #return(data)
  return(plot)
}
```


```{r}
plotbar(data = dafnia, y = "Feeding")
plotbar(data = dafnia, y = "Survival")
```


